---
title: "project"
author: "Erin Lyons, Yujian Tang, Norma Techarukpong, Ryan Thomas, Jason Yu"
date: "April 3, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(MASS)
library(ggplot2)
library(purrr)
library(leaps)
library(glmnet)
```

#Data Cleaning
```{r}
#Uploading Dataset
CPS.data <- read.csv("Chicago_Public_Schools_-_Progress_Report_Cards__2011-2012_.csv")

#Cleaning Data
CPS.data$Link <- NULL
CPS.data$Phone.Number <- NULL
CPS.data$State <- NULL
CPS.data$Street.Address <- NULL
CPS.data$Location <- NULL
CPS.data$City <- NULL
CPS.data$RCDTS.Code <- NULL
```

```{r}
#Removing NDA values from variables and changing the values to NA
for(i in names(CPS.data)) {
  for(j in 1:nrow(CPS.data)) {
    if(isTRUE(CPS.data[j,i] == "NDA")) { CPS.data[j, i] = NA}
  }
}

#View
head(CPS.data)
```
```{r}
#Data that only describes the numeric variables
numeric.data <- CPS.data[, map_lgl(CPS.data, is.numeric)]
head(numeric.data)
```

```{r}
# Erin's section
```

```{r}
# Charles' section
```

```{r}
# Norma's section
```


#Ryan's Section
Ideas:

- Other Variables connection to classificaiton
- Are there certain ways to categorize elementary middle and high school?

```{r}
#Safety and Parents
ggplot(CPS.data) +
  geom_point(aes(x = Safety.Score, y = Parent.Environment.Score, col = Adequate.Yearly.Progress.Made.))

#Attendance
ggplot(filter(CPS.data, Average.Teacher.Attendance != 0)) +
  geom_point(aes(x = Average.Student.Attendance, y = Average.Teacher.Attendance, col = Adequate.Yearly.Progress.Made.))

#
```

```{r}
numeric.data
```

#Numerical Analysis
```{r}
#Correlation of all Variables
correlation.matrix <- cor(numeric.data, use = "complete.obs")
correlation.dataset <- as.data.frame(correlation.matrix)
most.correlation <- which(abs(correlation.matrix) < 1 & abs(correlation.matrix) > .70)

rows.list <- list()
columns.list <- list()
j <- 1
#For Loop for finding maximum correlations
for(i in most.correlation) {
  #Correlation
  row <- ceiling(i/ncol(numeric.data))
  column <- ifelse(i%%ncol(numeric.data) == 0, ncol(numeric.data), i%%ncol(numeric.data))
  rows.list[[j]] <- rownames(correlation.dataset[row,])
  columns.list[[j]] <- names(correlation.dataset)[column]
  j = j+1
}

row.column.names = cbind(rows.list, columns.list)

#New Dataframe To hold Rows/ Columns
row.column.dataframe <- data.frame(
  row = NULL,
  column = NULL
)
#putting values in dataframe
for(i in 1:length(rows.list)) {
  newRow = data.frame(row = rows.list[[i]], column = columns.list[[i]])
  row.column.dataframe <- rbind(row.column.dataframe, newRow)
}

#Changing variable types to character
row.column.dataframe$row <- as.character(row.column.dataframe$row)
row.column.dataframe$column <- as.character(row.column.dataframe$column)

#Removing Duplicates
for (i in 1:nrow(row.column.dataframe))
{
    row.column.dataframe[i, ] = sort(row.column.dataframe[i, ])
}
row.column.dataframe <- row.column.dataframe[!duplicated(row.column.dataframe),]

#Viewing Dataframe of Variables with the most Correlation
row.column.dataframe
```



```{r}
#For Loop For Creating Scatterplots of data with the Greatest Correlation
for(i in 1:nrow(row.column.dataframe)) {
  #Extract column and row variable
  row <- row.column.dataframe[[i, 1]]
  column <- row.column.dataframe[[i, 2]]
  #Plot
  print(ggplot(CPS.data) +
    geom_point(aes_string(x = row, y = column), color = "blue4") +
    geom_smooth(aes_string(x = row, y = column)) +
    theme_minimal() +
    ggtitle("Correlations"))
}
```

##How well can we predict student attendance?
Goal - determine best model between lasso regression, ridge regression, and least squares regression with variable selection.
```{r}
#Least Squares Model
fullLinMod.SA <- lm(Average.Student.Attendance~., data = numeric.data)
```

```{r}
#Backward Elimination
BE.linMod.SA = regsubsets(Average.Student.Attendance~., data= numeric.data, nvmax= 20, method = "backward")
summary.backward = summary(BE.linMod.SA)
plot(summary.backward$cp, pch = 19)
```

```{r}
step(fullLinMod.SA, direction = "backward")
```

```{r}
linMod.Backward = lm(Average.Student.Attendance ~ Safety.Score + Environment.Score + 
    Instruction.Score + Rate.of.Misconducts..per.100.students. + 
    Individualized.Education.Program.Compliance.Rate + ISAT.Exceeding.Math.. + 
    ISAT.Exceeding.Reading.. + ISAT.Value.Add.Math + College.Enrollment..number.of.students. + 
    General.Services.Route + Community.Area.Number, data = numeric.data)

plot(linMod.Backward$fitted.values, linMod.Backward$residuals)
abline(0,0)
```

```{r}
#Foward Selection
FW.linMod.SA <- regsubsets(Average.Student.Attendance~., data= numeric.data, nvmax= 20, method = "forward")
summary.forward = summary(BE.linMod.SA)
plot(summary.forward$cp, pch = 19)
```
```{r}
step(fullLinMod.SA, direction = "forward")
```

```{r}
linMod.Forward <- lm(formula = Average.Student.Attendance ~ School.ID + ZIP.Code + 
    Safety.Score + Environment.Score + Instruction.Score + Rate.of.Misconducts..per.100.students. + 
    Average.Teacher.Attendance + Individualized.Education.Program.Compliance.Rate + 
    ISAT.Exceeding.Math.. + ISAT.Exceeding.Reading.. + ISAT.Value.Add.Math + 
    ISAT.Value.Add.Read + College.Enrollment..number.of.students. + 
    General.Services.Route + X_COORDINATE + Y_COORDINATE + Latitude + 
    Longitude + Community.Area.Number + Ward + Police.District, 
    data = numeric.data)

plot(linMod.Forward$fitted.values, linMod.Forward$residuals)
abline(0, 0)
```


```{r}
#Lasso Regression
modMatrix <- model.matrix(Average.Student.Attendance~., data= numeric.data)
lassoMod = cv.glmnet(modMatrix, y = na.omit(numeric.data)$Average.Student.Attendance, alpha = 1, nfolds = 5)
plot(lassoMod)
```
```{r}
#Ridge Regression
ridgeMod = cv.glmnet(modMatrix, y = na.omit(numeric.data)$Average.Student.Attendance, alpha = 1, nfolds = 5)
plot(ridgeMod)
```

##Classification for School Grade-level
```{r}
#Function that removes collinearity for data
remove.collinearity <- function(data, c) {
  #Correlation of all Variables
  correlation.matrix <- cor(data, use = "complete.obs")
  correlation.dataset <- as.data.frame(correlation.matrix)
  most.correlation <- which(abs(correlation.matrix) < 1 & abs(correlation.matrix) > c)

  rows.list <- list()
  columns.list <- list()
  j <- 1
  #For Loop for finding maximum correlations
  for(i in most.correlation) {
    #Correlation
    row <- ceiling(i/ncol(data))
    column <- ifelse(i%%ncol(data) == 0, ncol(data), i%%ncol(data))
    rows.list[[j]] <- rownames(correlation.dataset[row,])
    columns.list[[j]] <- names(correlation.dataset)[column]
    j = j+1
  }

  row.column.names = cbind(rows.list, columns.list)

  #New Dataframe To hold Rows/ Columns
  row.column.dataframe <- data.frame(
    row = NULL,
    column = NULL
  )
  #putting values in dataframe
  for(i in 1:length(rows.list)) {
    newRow = data.frame(row = rows.list[[i]], column = columns.list[[i]])
    row.column.dataframe <- rbind(row.column.dataframe, newRow)
  }

  #Changing variable types to character
  row.column.dataframe$row <- as.character(row.column.dataframe$row)
  row.column.dataframe$column <- as.character(row.column.dataframe$column)

  #Removing Duplicates
  for (i in 1:nrow(row.column.dataframe))
  {
      row.column.dataframe[i, ] = sort(row.column.dataframe[i, ])
  }
  row.column.dataframe <- row.column.dataframe[!duplicated(row.column.dataframe),]

  #Dataframe of Variables with the most Correlation
  table <- row.column.dataframe
  
  #creates new dataframe without correlations above c
  new.data = data
  for(i in 1:nrow(table)) {
    new.data[,table[i,2]] <- NULL
  }
  return(new.data)
}
```

```{r}
newdata = numeric.data %>%
  select_if(Negate(is.integer)) %>%
  select_if(is.numeric)

newdata <- remove.collinearity(newdata, .1)
newdata
newdata <- cbind(numeric.data, CPS.data$Elementary..Middle..or.High.School)
newdata$Elementary..Middle..or.High.School <- newdata$`CPS.data$Elementary..Middle..or.High.School`
newdata$`CPS.data$Elementary..Middle..or.High.School`<- NULL

ldaMod <- lda(Elementary..Middle..or.High.School~., data = newdata)
```

```{r}

```



```{r}
# Jason's section
```